Enforced Invariant Confluence

### 2018-03-04 doc/paper_redux/interactive_decision_procedure.tex
More interactive_decision_procedure tidying.
Revised everything with new s_0, T, I definitions.

\section{Interactive \Iconfluence{} Decision Procedure}
Imagine we have an \Iclosed{} decision procedure, and let's revisit
\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a set of
transactions $T$, and an invariant $I$. We want to know if $T$ is
\Iconfluent{}. Assume we have already determined that $I(s_0)$. We can hand $O$
and $I$ to the \Iclosed{} decision procedure to see if $O$ is \Iclosed{}. If
the decision procedure says yes, then by \clmref{StateBasedSufficient}, we know
that $(O, T)$ is \Iconfluent{}.

However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
say no. \clmref{SufficientAndNecessary} tells us that this means one of two
things. Either (1) $T$ is not \Iconfluent{} or (2) $(O, T)$ is \Iconfluent{},
but there are some invariant satisfying states that are not reachable. If we
think harder about the set of reachable states, we may realize that all
reachable states have a non-negative value of $x$. Thus, we may modify our
invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0 \land x
\geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the decision
procedure, it will say yes, and we can conclude that $(O, T)$ is \Iconfluent{}.

\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}
  \caption{Interactive \Iconfluence{} Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsIconfluent}{$O$, $T$, $I$}
      \If {$\lnot I(s_0)$}
        \State \Return no
      \Else
        \State \Return \Call{Helper}{$O$, $T$, $I$, $\set{s_0}$, $\emptyset$}
      \EndIf
    \EndFunction

@@

    \comment{$R$ is a set of reachable states.}
    \comment{$NR$ is a set of non-reachable states.}
    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
      \If {closed $=$ unknown}
        \State \Return unknown
      \ElsIf {closed $=$ yes}
        \State \Return{yes}
      \Else
        \State Augment $R$ and $NR$ with random search and user input
        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
        \If{$s_1, s_2 \in R$}
          \State \Return no
        \Else
          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\newcommand{\IsIconfluent}{\textsc{IsIconfluent}}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}
\newcommand{\Helper}{\textsc{Helper}}

This example motivates \algoref{InteractiveDecisionProcedure}, a general
purpose interactive \Iconfluence{} decision procedure in which a user
iteratively refines the invariant until the decision procedure terminates.
Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
\IsIconfluent$(O, T, I)$ to determine if $(O, T)$ is \Iconfluent{}.

@@

\IsIconfluent{} takes in $O$, $T$, and $I$. Formally, $I$ is set of states
(e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an actual
implementation of \IsIconfluent{} would likely take in $I$ as a logical formula
(e.g. $xy \leq 0$). From now on, whenever we mention a (possibly infinite) set
of states, remember that we can always represent the set as a finite formula.

\IsIconfluent{} calls a helper function \Helper{} which---in addition to $O$,
$T$, and $I$---takes as arguments a set $R$ of reachable states and a set $NR$
of unreachable states. Like \IsIconfluent, \Helper$(O, T, I, R, NR)$ returns
whether $T$ is \Iconfluent. Intuitively, $NR$ helps remove unreachable states
from $I$ moving us closer to the case where $I \subseteq
\setst{e}{\reachable{e}}$.

\Helper{} first checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
cases to consider.
\begin{itemize}
  \item
    If \IsIclosed{} gets stuck and doesn't know whether or not $O$ is
    \iclosed{$(I - NR)$}, then \Helper{} is stuck too and is forced to return
    unknown.
  \item
    If \IsIclosed{} determines that $O$ is \iclosed{$(I - NR)$}, then by
    \clmref{StateBasedSufficient}, $T$ is \iconfluent{$(I - NR)$}, so
    $\setst{e}{\reachable{e}} \subseteq I - NR$. $I - NR \subseteq I$, so $T$
    is therefore also \Iconfluent{}. Thus, \Helper{} return yes.
  \item
    If \IsIclosed{} determines that $O$ is not \iclosed{$(I - NR)$}, then we
    have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
    and $s_2$ are reachable or unreachable. We can do so in two ways.

@@

    Second, we can prompt the user to tell us whether or not the states are
    reachable or unreachable. In addition to labelling $s_1$ and $s_2$ as
    reachable or not, the user can also refine $I$ by augmenting $NR$. Taking
    \exampleref{StateBasedNotNecessary} as an example, the user might look
    through the existing reachable and unreachable states and have an aha
    moment realizing that every point with a negative $x$-coordinate is
    unreachable. They can use this knowledge to add all points with negative
    $x$-coordinates to $NR$.

    After $s_1$ and $s_2$ have been labelled reachable or not, we continue. If
    both $s_1$ and $s_2$ are reachable, then so is $s_1 \join s_2$, but $\lnot
    I(s_1 \join s_2)$. Thus, $(O, T)$ is not \Iconfluent{}, so \Helper{}
    returns no.  Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we
    recurse.
\end{itemize}

An example of how to use \algoref{InteractiveDecisionProcedure} on
\exampleref{StateBasedNotNecessary} is given in
\figref{DecisionProcedureExample}.

@@

\todo{Describe what happens when user input is incorrect.}

>>>>

\section{Interactive Invariant-Confluence Decision Procedure}
Imagine we have an \Iclosed{} decision procedure, and let's revisit
\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a start
state $s_0$, set of transactions $T$, and an invariant $I$. We want to know if
$O$ is \sTIconfluent{}. We can hand $O$ and $I$ to the \Iclosed{} decision
procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes and
we can conclude that $I(s_0)$, then by \clmref{StateBasedSufficient}, we know
that $O$ is \sTIconfluent{}.

However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
say no. \clmref{SufficientAndNecessary} tells us that this means one of two
things. Either (1) $O$ is \emph{not} \sTIconfluent{} or (2) $O$ is
\sTIconfluent{}, but there are some invariant satisfying states that are not
reachable. If we think harder about the set of reachable states, we may realize
that all reachable states have a non-negative value of $x$. Thus, we may modify
our invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0
\land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the
decision procedure, it will say yes. Moreover, $I(s_0)$, so we can conclude
that $O$ is \sTIconfluent{}.

\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
\begin{algorithm}
  \caption{Interactive Invariant-Confluence Decision Procedure}%
  \algolabel{InteractiveDecisionProcedure}
  \begin{algorithmic}
    \Function{IsIconfluent}{$O$, $s_0$, $T$, $I$}
      \State
        \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
    \EndFunction

@@

    \comment{$R$ is a set of $\sTIreachable{}$ states.}
    \comment{$NR$ is a set of $\sTIunreachable{}$ states.}
    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
      \If {$\lnot(I - NR)(s_0)$}
        \State \Return no
      \EndIf
      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
      \If {closed $=$ unknown}
        \State \Return unknown
      \ElsIf {closed $=$ yes}
        \State \Return{yes}
      \Else
        \State Augment $R$ and $NR$ with random search and user input
        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
        \If{$s_1, s_2 \in R$}
          \State \Return no
        \Else
          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\todo{Rename IsIconfluent to IsSTIConfluent}
\newcommand{\IsIconfluent}{\textsc{IsIconfluent}}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}
\newcommand{\Helper}{\textsc{Helper}}

This example motivates \algoref{InteractiveDecisionProcedure}, a general
purpose interactive invariant-confluence decision procedure in which a user
iteratively refines the invariant until the decision procedure terminates.
Given an object $O$, a start state $s_0$, a set of transactions $T$, and an
invariant $I$, we call \IsIconfluent$(O, s_0, T, I)$ to determine if $O$ is
\sTIconfluent{}. Now, we describe \algoref{InteractiveDecisionProcedure}. An
example of how to use \algoref{InteractiveDecisionProcedure} on
\exampleref{StateBasedNotNecessary} is given in
\figref{DecisionProcedureExample}.

@@

\IsIconfluent{} takes in $O$, $s_0$, $T$, and $I$. Formally, $I$ is set of
states (e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an
actual implementation of \IsIconfluent{} would likely take in $I$ as a logical
formula (e.g. $xy \leq 0$). From now on, whenever we mention a (possibly
infinite) set of states, remember that we can always represent the set as a
finite formula.

\IsIconfluent{} calls a helper function \Helper{} which---in addition to $O$,
$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
and a set $NR$ of $\sTIunreachable{}$ states. Like \IsIconfluent, \Helper$(O,
s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
$NR$ are correct). Intuitively, $NR$ helps remove unreachable states from $I$
moving us closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.

\Helper{} first checks to see if $(I - NR)(s_0)$. If it is not, then \Helper{}
returns no. If $NR$ only contains unreachable states, then it doesn't contain
$s_0$, so $s_0 \in I - NR \iff s_0 \in I$. Thus, if $\lnot (I - NR)(s_0)$, then
$\lnot I(s_0)$, so $O$ is not \sTIconfluent.

Next, \Helper{} checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
cases to consider.
\begin{itemize}
  \item
    If \IsIclosed{} gets stuck and doesn't know whether or not $O$ is
    \iclosed{$(I - NR)$}, then \Helper{} is stuck too and is forced to return
    unknown.

  \item
    If \IsIclosed{} determines that $O$ is \iclosed{$(I - NR)$}, then by
    \clmref{StateBasedSufficient}, $O$ is \sticonfluent{s_0}{T}{(I -
    NR)}, so
    $
      \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
        \subseteq I - NR
        \subseteq I
    $.
    If $NR$ only contains unreachable states, then
    $
      \setst{s}{\sTIreachable{s}} = \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
    $.
    Thus,
    $
      \setst{s}{\sTIreachable{s}} \subseteq I
    $.
    Thus, $O$ \sTIconfluent{}, so \Helper{} return yes.

  \item
    If \IsIclosed{} determines that $O$ is not \iclosed{$(I - NR)$}, then we
    have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
    and $s_2$ are $\sTIreachable{}$ or $\sTIunreachable{}$. We can do so in two
    ways.

@@

    Second, we can prompt the user to tell us whether or not the states are
    $\sTIreachable{}$ or $\sTIunreachable{}$. In addition to labelling $s_1$
    and $s_2$ as $\sTIreachable{}$ or not, the user can also refine $I$ by
    augmenting $NR$. Taking \exampleref{StateBasedNotNecessary} as an example,
    the user might look through the existing $\sTIreachable{}$ and
    $\sTIunreachable{}$ states and have an aha moment realizing that every
    point with a negative $x$-coordinate is $\sTIunreachable{}$. They can use
    this knowledge to add all points with negative $x$-coordinates to $NR$.

    After $s_1$ and $s_2$ have been labelled $\sTIreachable{}$ or not, we
    continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1
    \join s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{},
    so \Helper{} returns no. Otherwise, one of $s_1$ and $s_2$ is
    $\sTIunreachable{}$. Now, we recurse.
\end{itemize}

@@

% \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
% procedure. It requires that a user classify states as reachable or unreachable.
% What happens if a user incorrectly classifies a state? Given a counterexample
% $s_1, s_2$, if a user incorrectly says that one of the unreachable points is
% reachable, then the $s_1, s_2 \in R$ check in \Helper{} might be
% erroneously met. In turn, \IsIconfluent{} might erroneously return no. This is
% inconvenient, but not the end of the world.
% Moreover, we can modify our decision procedure so that \Helper{}
% user provide a \emph{proof} of reachability whenever they label as: a series of transactions and joins that result in the

### 2018-03-04 doc/paper_redux/new_statebased_sufficient.tex
Revised everything with new s_0, T, I definitions.

\section{New Sufficient State-Based Conditions}
\todo{%
  Revise everything with new notation of $(O, T)$ \Iconfluent{} instead of $T$
  \Iconfluent{}.
}

\algoref{InteractiveDecisionProcedure} relies on a user to provide information
about reachable and unreachable states. In this section, we outline a new
sufficient condition for state-based \Iconfluence{} that does not require user
interaction and that covers some cases not covered by the sufficient conditions
in \clmref{StateBasedSufficient}. To prove that the new condition is
sufficient, we introduce the concepts of \dIconfluence{} and \dIreduction{}.

\begin{definition}
  A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s_0) \ldots))$ is an
  expression that does not contain a join.
\end{definition}

\begin{definition}
  $T$ is \defword{\dIconfluent{}} if for every pair $e_1$ and $e_2$ of
  join-free expressions, $\Irec{e_1}$ and $\Irec{e_2}$ implies that $I(e_1
  \join e_2)$.
\end{definition}

Intuitively, \dIconfluence{} is the same thing as \Iconfluence{} with
the added restriction that subexpressions are join-free.

\begin{definition}
  $T$ is \defword{\dIreducible{}} if for every pair $e_1$ and $e_2$ of
  join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some
  join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1
  \join e_2)$.
\end{definition}

Intuitively, \dIreduction{} allows us to eliminate join operators by
substituting $e_1 \join e_2$ (which contains 1 join) with $e_3$ (which does not
contain any joins).

% \input{diconfluence_direduction.fig}
%
% We can also interpret \dIconfluence{} and \dIreduction{} using the graph-based
% model of \Iconfluence{}, as illustrated in
% \figref{diconfluence_and_direduction}. Intuitively, \dIconfluence{} is the
% property that any two divergent join-free paths that satisfy $I$ can be joined
% together to form a new state that satisfies $I$. \dIreduction{} is the property
% that the join of any two divergent join-free paths that satisfy $I$ can be
% replaced with a single join-free path that satisfies $I$. Both conditions
% involve diamond diagrams, hence the names \dIconfluence{} and \dIreduction{}.

\begin{claim}\clmlabel{DiconfluenceDireducibleImpliesIconfluence}
  If $T$ is \dIconfluent{} and \dIreducible, then it is \Iconfluent.
\end{claim}
\begin{elidableproof}
  First, we prove that for all $e$, if $\Irec{e}$ then there exists a join-free
  $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty
  straightforward structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case 1: $e = s_0$.}
      Trivially, $e' = s_0.$

    \item \textbf{Case 2: $e = t(e_1)$.}
      $\Irec{t(e_1)}$, so $\Irec{e_1}$. Thus, by the inductive hypothesis,
      there exists a join-free $e_1'$ such that $\Irec{e_1'}$ and $eval(e_1) =
      eval(e_1')$. Now, let $e' = t(e_1')$ which is clearly join-free.
      \[
        eval(e')
          = eval(t(e_1'))
          = t(eval(e_1'))
          = t(eval(e_1))
          = eval(t(e_1))
          = eval(e)
      \]
      This also implies that $\Irec{e'}$.

    \item \textbf{Case 3: $e = e_1 \join e_2$.}
      $\Irec{e_1 \join e_2}$, so $\Irec{e_1}$ and $\Irec{e_2}$. Thus, by the
      inductive hypothesis, there exist join-free expressions $e_1'$ and $e_2'$
      where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$,
      and $\Irec{e_2'}$. Let $e'' = e_1' \join e_2'$
      \begin{align*}
        eval(e'')
        &= eval(e_1' \join e_2') \\
        &= eval(e_1') \join eval(e_2') \\
        &= eval(e_1) \join eval(e_2) \\
        &= eval(e_1 \join e_2) \\
        &= eval(e)
      \end{align*}
      This implies that $\Irec{e''}$. Because $T$ is \dIreducible, there exists
      a join-free $e'$ such that $\Irec{e'}$ and $eval(e') = eval(e'') =
      eval(e)$.
  \end{itemize}

  Next, consider arbitrary expressions $e_1$ and $e_2$ where $\Irec{e_1}$ and
  $\Irec{e_2}$. To show that $T$ is \Iconfluent, we must show that $I(e_1 \join
  e_2)$. Using our little lemma above, we know there exists join-free $e_1'$
  and $e_2'$ where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') =
  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join
  e_2')$, and
  \[
    eval(e_1' \join e_2')
      = eval(e_1') \join eval(e_2')
      = eval(e_1) \join eval(e_2)
      = eval(e_1 \join e_2)
  \]
  Thus, $I(e_1 \join e_2)$.
\end{elidableproof}

\begin{claim}
  Consider a state-based object $(S, s_0, \join)$, a set of transactions $T$,
  and an invariant $I$. $T$ is \Iconfluent{} if the following criteria are met:
  \begin{enumerate}
    \item
      $(S, \join)$ is a semilattice.

@@

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  \newcommand{\bart}[1]{\overline{t_{#1}}}
  \newcommand{\baru}[1]{\overline{u_{#1}}}

  To show that $T$ is \Iconfluent{}, it suffices to show that $T$ is
  \dIconfluent{} and \dIreducible{} (by
  \clmref{DiconfluenceDireducibleImpliesIconfluence}). We begin by showing
  that $T$ is \dIconfluent{}. Let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free expressions that recursively satisfy $I$. For ease
  of notation, let
  \[
    \bart{i} = t_i(\ldots(t_1(s_0))\ldots)
    \quad\text{and}\quad
    \baru{i} = u_j(\ldots(u_1(s_0))\ldots)
  \]
  We prove by strong induction on $m \in \nats$ that if $m = i + j$ where $0
  \leq i \leq n$ and $0 \leq j \leq m$, then $\Irec{\bart{i}(\baru{j}(s_0))}$.
  From this, it's immediate that $\Irec{\bart{n}(\baru{m}(s_0))}$, so
  $\Irec{e_1 \join e_2}$, which proves that $T$ is \dIconfluent{}.
  \begin{itemize}
    \item \textbf{Case $m = 0$.}
      $i = j = 0$, so $\bart{0}(\baru{0}(s_0)) = s_0$. $\Irec{s_0}$ because
      $\Irec{e_1}$.

    \item \textbf{Case $m = 1$.}
      Without loss of generality, assume $i = 1$ and $j = 0$. Then,
      \[
        \bart{1}(\baru{0}(s_0))
          = t_1(s_0)
      \]
      $\Irec{t_1(s_0)}$ because $\Irec{e_1}$.

@@

  Next, we show that $T$ is \dIreducible{}. Again let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free expressions that recursively satisfy $I$. We just
  proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also, $eval(e_1 \join e_2) =
  eval(\bart{n}(\baru{m}(s_0)))$. $\bart{n}(\baru{m}(s_0))$ is a join-free
  expression, so $T$ is \dIreducible{}.
\end{elidableproof}

>>>>

\section{New Sufficient State-Based Conditions}
\algoref{InteractiveDecisionProcedure} relies on a user to provide information
about reachable and unreachable states. In this section, we outline a new
sufficient condition for state-based invariant-confluence that does not require
user interaction and that covers some cases not covered by the sufficient
conditions in \clmref{StateBasedSufficient}. To prove that the new condition is
sufficient, we introduce the concepts of \dsTIconfluence{} and
\dsTIreduction{}.

\begin{definition}
  A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s) \ldots))$ is an
  expression that does not contain a join.
\end{definition}

\begin{definition}
  $O$ is \defword{\dsTIconfluent{}} if (1) $I(s_0)$ and (2) for every pair
  $e_1$ and $e_2$ of join-free $\sTIreachable{}$ expressions, $\Irec{e_1}$ and
  $\Irec{e_2}$ implies that $I(e_1 \join e_2)$.
\end{definition}

Intuitively, \dsTIconfluence{} is the same thing as the second condition in
\clmref{StateBasedTwoIconfluenceDefs} with the added restriction that
$e_1$ and $e_2$ are join-free.

\begin{definition}
  $O$ is \defword{\dsTIreducible{}} if for every pair $e_1$ and $e_2$ of
  join-free $\sTIreachable{}$ expressions, $\Irec{e_1 \join e_2}$ implies that
  there exists some join-free $\sTIreachable{}$ expression $e_3$ such that
  $\Irec{e_3}$ and $eval(e3) = eval(e1 \join e_2)$.
\end{definition}

Intuitively, \dsTIreduction{} allows us to eliminate join operators by
substituting $e_1 \join e_2$ (which contains 1 join) with $e_3$ (which does not
contain any joins).

\begin{claim}\clmlabel{DiconfluenceDireducibleImpliesIconfluence}
  If $O$ is \dsTIconfluent{} and \dsTIreducible, then $O$ is \sTIconfluent.
\end{claim}
\begin{elidableproof}
  First, we prove that for all expressions $e$, if $\sTIreachable{e}$ and
  $\Irec{e}$ then there exists a join-free $\sTIreachable{}$ $e'$ where
  $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty straightforward
  structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case 1: $e = s_0$.}
      $s_0$ is join-free, so we can trivially let $e' = s_0$.

    \item \textbf{Case 2: $e = t(e_1)$.}
      $\sTIreachable{t(e_1)}$ and $\Irec{t(e_1)}$, so $\sTIreachable{e_1}$ and
      $\Irec{e_1}$. Thus, by the inductive hypothesis, there exists a join-free
      $\sTIreachable{}$ $e_1'$ such that $\Irec{e_1'}$ and $eval(e_1) =
      eval(e_1')$. Now, let $e' = t(e_1')$ which is clearly join-free.
      \[
        eval(e')
          = eval(t(e_1'))
          = t(eval(e_1'))
          = t(eval(e_1))
          = eval(t(e_1))
          = eval(e)
      \]
      This also implies that $\Irec{e'}$.

    \item \textbf{Case 3: $e = e_1 \join e_2$.}
      $\sTIreachable{e_1 \join e_2}$ and $\Irec{e_1 \join e_2}$, so
      $\sTIreachable{e_1}$, $\sTIreachable{e_2}$, $\Irec{e_1}$ and
      $\Irec{e_2}$. Thus, by the inductive hypothesis, there exist join-free
      $\sTIreachable{}$ expressions $e_1'$ and $e_2'$ where $eval(e_1') =
      eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$, and $\Irec{e_2'}$.
      Let $e'' = e_1' \join e_2'$
      \begin{align*}
        eval(e'')
        &= eval(e_1' \join e_2') \\
        &= eval(e_1') \join eval(e_2') \\
        &= eval(e_1) \join eval(e_2) \\
        &= eval(e_1 \join e_2) \\
        &= eval(e)
      \end{align*}
      This implies that $\Irec{e''}$. Because $O$ is \dsTIreducible, there
      exists a join-free $\sTIreachable{}$ $e'$ such that $\Irec{e'}$ and
      $eval(e') = eval(e'') = eval(e)$.
  \end{itemize}

  Next, consider arbitrary $\sTIreachable{}$ expressions $e_1$ and $e_2$ where
  $\Irec{e_1}$ and $\Irec{e_2}$. To show that $T$ is \sTIconfluent, we must
  show that $I(e_1 \join e_2)$. Using our little lemma above, we know there
  exists join-free $\sTIreachable{}$ $e_1'$ and $e_2'$ where $eval(e_1') =
  eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$, and $\Irec{e_2'}$.
  Because $T$ is \dsTIconfluent, $I(e_1' \join e_2')$, and
  \[
    eval(e_1' \join e_2')
      = eval(e_1') \join eval(e_2')
      = eval(e_1) \join eval(e_2)
      = eval(e_1 \join e_2)
  \]
  Thus, $I(e_1 \join e_2)$.
\end{elidableproof}

\begin{claim}
  Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
  transactions $T$, and an invariant $I$. $T$ is \sTIconfluent{} if the
  following criteria are met:
  \begin{enumerate}
    \item
      $O$ is a semilattice.

@@

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.

    \item
      $I(s_0)$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  \newcommand{\bart}[1]{\overline{t_{#1}}}
  \newcommand{\baru}[1]{\overline{u_{#1}}}

  To show that $T$ is \sTIconfluent{}, it suffices to show that $T$ is
  \dsTIconfluent{} and \dsTIreducible{} (by
  \clmref{DiconfluenceDireducibleImpliesIconfluence}). We begin by showing that
  $T$ is \dsTIconfluent{}. Let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free $\sTIreachable{}$ expressions that recursively
  satisfy $I$. For ease of notation, let
  \[
    \bart{i} = t_i(\ldots(t_1(s_0))\ldots)
    \quad\text{and}\quad
    \baru{i} = u_j(\ldots(u_1(s_0))\ldots)
  \]
  We prove by strong induction on $m \in \nats$ that if $m = i + j$ where $0
  \leq i \leq n$ and $0 \leq j \leq m$, then $\Irec{\bart{i}(\baru{j}(s_0))}$.
  From this, it's immediate that $\Irec{\bart{n}(\baru{m}(s_0))}$, so
  $\Irec{e_1 \join e_2}$, which proves that $T$ is \dsTIconfluent{}.
  \begin{itemize}
    \item \textbf{Case $m = 0$.}
      $i = j = 0$, so $\bart{0}(\baru{0}(s_0)) = s_0$. $\Irec{s_0}$ by
      assumption.

    \item \textbf{Case $m = 1$.}
      Without loss of generality, assume $i = 1$ and $j = 0$. Then,
      $
        \bart{1}(\baru{0}(s_0))
          = t_1(s_0)
      $.
      $\Irec{t_1(s_0)}$ because $\Irec{e_1}$.

@@

  Next, we show that $T$ is \dsTIreducible{}. Again let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free $\sTIreachable{}$ expressions that recursively
  satisfy $I$. We just proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also,
  $eval(e_1 \join e_2) = eval(\bart{n}(\baru{m}(s_0)))$.
  $\bart{n}(\baru{m}(s_0))$ is a join-free $\sTIreachable{}$ expression, so $T$
  is \dsTIreducible{}.
\end{elidableproof}
